---
title: "Ejercicios SAND capítulo 1 y 2"
output: html_notebook
---

```{r setup}
library(sand)
library(igraph)
```


#  Creating Network Graphs

## Undirected and Directed Graphs
### Creating undirected graphs

```{r}

g <- graph_from_literal(1-2, 1-3, 2-3, 2-4, 3-5, 4-5, 4-6, 4-7, 5-6, 6-7)
```

```{r}
g1 <- graph.formula(1-2, 1-3, 2-3, 2-4, 3-5, 4-5, 4-6, 4-7, 5-6, 6-7, 8-7)
```

### Retrieving vertices

```{r}
V(g1)
```



### Retrieving edges

```{r}
E(g1)
```

### Printing a summary

```{r}
print_all(g1)
```


### PLotting the graph

```{r}
plot(g1)
```



### Creating directed graphs

```{r}
dg <- graph_from_literal(1-+2, 1-+3, 2++3)
dg1 <- graph.formula(1-+2, 1-+3, 2++3)
plot(dg)

```


### Creating undirected graphs with given names

```{r}
dg2 <- graph_from_literal(Sam-+Mary, Sam-+Tom, Mary++Tom)
print_all(dg2)
```

### Rename the vertices

```{r}
V(dg)$name <- c("Sam", "Mary", "Tom")
```



## Representations for Graphs
### Adjacency list representation

```{r}
dg.adjacency <- get.adjacency(dg)
dg2 <- graph.adjlist(dg.adjacency + 1)
```

### Edge list representation

```{r}
E(dg2)
```

```{r}
dg2.edgelist <- get.edgelist(dg2)
dg2.edgelist
```

```{r}
graph.edgelist(dg2.edgelist)
```

### Adjacency representation

```{r}
as_adjacency_matrix(g)
graph.adjacency(g)
```


## Operations on Graphs
### Obtaining a subgraph

```{r}
h <- induced_subgraph(g, 1:5)
print_all(h)
```


### Obtaining a subgraph by the minus operator

```{r}
h <- g - vertices(c(6,7))
```

### Recovering g from h by the plus operator

```{r}
h <- h + vertices(c(6,7))
g <- h + edges(c(4,6),c(4,7),c(5,6),c(6,7))
print_all(g)
```


### Applying union to obatin a graph

```{r}
h1 <- h
h2 <- graph_from_literal(4-6, 4-7, 5-6, 6-7)
g <- union(h1,h2)
print_all(g)
```


## Decorating Network-Graphs
### Vertex, Edge, and Graph Attributes
### Listing vertices from dg

```{r}
V(dg)$name
```


# Adding attributes to vertices

```{r}
V(dg)$gender <- c("M","F","M")
```

# Associating te color red with vertices

```{r}
V(g)$color <- "red"
plot(g)
```


# Verifying if a graph is weighted

```{r}
is_weighted(g)
wg <- g
E(wg)$weight <- runif(ecount(wg))
is_weighted(wg)

```


# Linking a name to a graph

```{r}
g$name <- "Toy Graph"
```

## Using data frames
### Importing information from vertices and edges data frames

```{r}
g.lazega <- graph_from_data_frame(
  elist.lazega,
  directed="FALSE",
  vertices=v.attr.lazega
)
g.lazega$name <- "Lazega Lawyers"
plot(g.lazega)
```

# Network information

```{r}
vcount(g.lazega)
ecount(g.lazega)
```

# Listing vertex attributes

```{r}
vertex_attr_names(g.lazega)
```

# Talking about graphs
## Basic graph concepts
### Veerifying if a graph is simple

```{r}
is_simple(g)
```

# Converting g to a multi-graph (loops and multi-edges)

```{r}
mg <- g + edge(2,3)
print_all(mg)
is_simple(mg)
```

# Converting multi-graph to a weighted graph

```{r}
E(mg)$weight <- 1
wg2 <- simplify(mg)
is_simple(wg2)
print_all(wg2)
E(wg2)$weight
```


# Listing neighbors

```{r}
neighbors(g,5)
```

# Calculating the degree of vertices (incident edges) on undirected graph

```{r}
degree(g)
```

# Calculating the degree of vertices (incident edges) on directed graph

```{r}
degree(dg, mode="in")
degree(dg, mode="out")
```

# Check the connectivity in a undirected graph

```{r}
is_connected(g)
```

# Listing components

```{r}
clusters(g)
```

# Check the connectivity in a directed graph

```{r}
is_connected(dg,mode="weak")
is_connected(dg,mode="strong")
```


# Calculating diameter of the graph

```{r}
diameter(g, weights=NA)
```

# Special Types of Graphs

```{r}
g.full <- make_full_graph(7)
g.ring <- make_ring(7)
g.tree <- make_tree(7, children=2, mode="undirected")
g.star <- make_star(7, mode="undirected")
par(mfrow=c(2, 2), mai = c(0.2, 0.2, 0.2, 0.2))
plot(g.full)
plot(g.ring)
plot(g.tree)
plot(g.star)
```


# Veryfing is a graph is a DAG

```{r}
is_dag(dg)
```

# Creating a bipartite graph

```{r}
g.bip <- graph_from_literal(
  actor1:actor2:actor3,
  movie1:movie2, actor1:actor2 - movie1,
  actor2:actor3 - movie2
)
V(g.bip)$type <- grepl("^movie", V(g.bip)$name)
print_all(g.bip, v=T)
```

# Creating a projection

```{r}
proj <- bipartite_projection(g.bip)
print_all(proj[[1]])
print_all(proj[[2]])
```
